---
title: "Montgomery County Traffic Incidents"
author: "Michael Metzler"
institute: "Montgomery College"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: false
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = FALSE, fig.width = 12, fig.height = 6, dpi = 300)
library(tidyverse)
library(lubridate)
library(leaflet)
incidents <- feather::read_feather("incidents.feather")
drivers <- feather::read_feather("drivers.feather")
non_motorists <- feather::read_feather("non_motorists.feather")
single_color <- "#3182bd"
basic_theme <- theme(
  plot.title = element_text(size = 24, vjust = .5, hjust = .5),
  axis.title = element_text(size = 16, hjust = .5, vjust = .5),
  panel.background = element_rect(fill = "white"),
  legend.position = "none",
  panel.grid.major.y = element_line(color = "grey75", size = .1, linetype = "solid")
)
```

## Automated Crash Reporting System

- 3 Datasets (Incidents, Drivers, Non-Motorists)
- `r nrow(incidents)` incidents since January 1st 2015
- `r nrow(drivers)` drivers
- `r nrow(non_motorists)` non-motorists
- The data is mostly categorical or descriptions
- Variety of NA encodings in the data (N/A, NA, empty strings)

---
class: center, top
## Data Cleaning and Wrangling

class: left, middle
- Identifyied data points with geolocation not in the county
- Use One-Hot encoding to create dummy variables for certain categorical variables
- Combine the incidents, drivers, and non-motorists tables into a single combined table with count data per incident

---
## Types of Incidents

```{r incidenttypes}
incidents %>%
  group_by(acrs_report_type) %>%
  count() %>%
  rename(`Report Type` = acrs_report_type, Incidents = n) %>%
  knitr::kable(format = "html")
```

---
class: center, middle

```{r}
incidents %>%
  select(incident_hour, incident_minute) %>%
  mutate(incident_time = 60 * incident_hour + incident_minute) %>%
  ggplot() +
  stat_count(aes(x = incident_time), fill = single_color) +
  scale_x_continuous(
    name = "Time",
    breaks = seq.int(from = 0, to = 24 * 60, by = 120),
    labels = c(paste0(seq.int(from = 0, to = 22, by = 2), ":00"),"0:00")
  ) +
  ggtitle("Crashes by Time of Day") +
  basic_theme
```

---

```{r}
incidents %>%
  select(incident_weekday) %>%
  ggplot() +
  stat_count(aes(x = incident_weekday), fill = single_color) +
  scale_x_continuous(
    name = "Day of Week",
    breaks = 1:7,
    labels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thurday", "Friday", "Saturday")
  ) +
  ggtitle("Crashes by Day of Week") +
  basic_theme
```

---

```{r}
incidents %>%
  select(incident_year) %>%
  ggplot() +
  stat_count(aes(x = incident_year), fill = single_color) +
  ggtitle("Crashes by Year") +
  basic_theme
```
---
class: center
# Number of Crashes by Date
```{r}
incidents %>%
  filter(incident_year < 2019) %>%
  select(incident_month, incident_day) %>%
  mutate(incident_date = paste0(incident_month, "/", incident_day)) %>%
  count(incident_date) %>%
  arrange(desc(n)) %>%
  rename(Date = incident_date, `Number of Crashes` = n) %>%
  DT::datatable()
```

---
class: center
# Weather

```{r incidentsweather}
incidents %>%
  group_by(weather) %>%
  count() %>%
  arrange(desc(n)) %>%
  knitr::kable(format="html")
```

---

class: center

```{r}
incidents %>%
  filter(driver_substance_abuse_alcohol_present | driver_substance_abuse_alcohol_contributed) %>%
  select(incident_hour, incident_minute) %>%
  mutate(incident_time = 60 * incident_hour + incident_minute) %>%
  ggplot() +
  stat_count(aes(x = incident_time), fill = single_color) +
  scale_x_continuous(
    name = "Time",
    breaks = seq.int(from = 0, to = 24 * 60, by = 120),
    labels = paste0(seq.int(from = 0, to = 24, by = 2), ":00")
  ) +
  ggtitle("Crashes Involving Alcohol by Time of Day") +
  basic_theme
```

---

```{r}
incidents %>%
  filter(driver_substance_abuse_alcohol_present | driver_substance_abuse_alcohol_contributed) %>%
  select(incident_weekday) %>%
  ggplot() +
  stat_count(aes(x = incident_weekday), fill = single_color) +
  scale_x_continuous(
    name = "Day of Week",
    breaks = 1:7,
    labels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thurday", "Friday", "Saturday")
  ) +
  ggtitle("Crashes Involving Alcohol by Day of Week") +
  basic_theme
```

---

class: center

```{r}
ordered_levels <- drivers %>%
  count(injury_severity) %>%
  arrange(desc(n)) %>%
  pull(injury_severity)
drivers %>%
  mutate(`Injury Severity` = factor(injury_severity, ordered_levels)) %>%
  ggplot() +
  stat_count(aes(x = `Injury Severity`), fill = single_color) +
  ggtitle("Severity of Driver Injuries") +
  basic_theme
```
---
class: center
```{r}
ordered_levels <- drivers %>%
  count(vehicle_damage_extent) %>%
  arrange(desc(n)) %>%
  pull(vehicle_damage_extent)
drivers %>%
  mutate(`Vehicle Damage` = factor(vehicle_damage_extent, ordered_levels)) %>%
  ggplot() +
  stat_count(aes(x = `Vehicle Damage`), fill = single_color) +
  ggtitle("Severity of Vehicle Damage") +
  basic_theme
```
---
class: center
```{r}
drivers %>%
  ggplot() +
  stat_count(aes(x = speed_limit, fill = ), fill = single_color) +
  ggtitle("Speed Limits") +
  basic_theme
```
---
class: center
```{r}
ordered_injury <- drivers %>%
  count(injury_severity) %>%
  arrange(desc(n)) %>%
  pull(injury_severity)
drivers %>%
  mutate(
    speed_limit = factor(c("LOW", "MEDIUM", "HIGH")[1 + findInterval(speed_limit, c(31, 46))], levels = c("LOW", "MEDIUM", "HIGH")),
    injury_severity = factor(injury_severity, ordered_injury)
  ) %>%
  group_by(speed_limit, injury_severity) %>%
  tally() %>%
  rename(`Speed Limit` = speed_limit, `Injury Severity` = injury_severity) %>%
  ggplot() +
  geom_bar(aes(x = `Injury Severity`, weight = n, fill = `Speed Limit`), position = "fill") +
  basic_theme +
  theme(legend.position = "top") +
  ggtitle("Injuries and Speed Limits") + 
  ylab("Proportion")
```
---
```{r}
ordered_levels <- drivers %>%
  count(vehicle_body_type) %>%
  arrange(desc(n)) %>%
  pull(vehicle_body_type)
drivers %>%
  mutate(`Body Type` = factor(as.character(fct_collapse(vehicle_body_type, OTHER = ordered_levels[-(1:5)])), c(ordered_levels[1:5], "OTHER"))) %>%
  ggplot() +
  stat_count(aes(x = `Body Type`), fill = single_color) +
  ggtitle("Vehicle Body Types") +
  basic_theme
```
---
```{r}
ordered_levels <- non_motorists %>%
  count(pedestrian_type) %>%
  arrange(desc(n)) %>%
  pull(pedestrian_type)
non_motorists %>%
  mutate(pedestrian_type = factor(pedestrian_type, ordered_levels)) %>%
  ggplot() +
  stat_count(aes(x = pedestrian_type), fill = single_color) +
  basic_theme +
  ggtitle("Types of Non-Motorists")
```
---
```{r}
pal <- colorFactor(c("blue", "red"), domain = c(0, 1))
non_motorists %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(lng = ~longitude, lat = ~latitude, radius = 3, stroke = TRUE, weight = 2, opacity = 1, color = ~ pal(not_in_county))
```